---
title: "Learning Python: Part 2"
author: ~
date: '2019-06-17'
slug: learning-python-2Rmd
categories: ['Python']
tags: ['Python', 'Pandas']
description: 'Part two of translating data tasks done in R but with Python.'
topics: ['Python']
---

```{r setup, echo=FALSE}
# Set up options for rmarkdown chunks
knitr::opts_chunk$set(warning = FALSE, message = FALSE, comment = "#")
```

# Learning Python

This post is a continuation of a learning python. In this post, I will extend the last post and focus on tidyverse equivalents to:

1. Mutate columns
2. Summarize data (Groupby)
3. Pipe functions (**%>%**)

# Setting up Python in Rmarkdown

```{r setup-python, echo=FALSE}
library(reticulate)
use_virtualenv("r-reticulate")
```

## Load Modules and Data
```{python load}
import pandas as pd
import numpy as np

url = 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-28/winemag-data-130k-v2.csv'
dat = pd.read_csv(url, index_col=0)
```

## Settings for better output 

This blog is built through the **blogdown** package and uses Rmarkdown files that render to HTML documents. Rmarkdown files best support R output but I do not know how to best format python outputs. I get inconsistent syntax highlighting. 

```{python options}
# Stop Python from truncating number of columns
pd.set_option('display.max_columns', None)
pd.set_option('display.width', 1000)
pd.set_option('precision', 2)
```

```{r echo=FALSE, eval=FALSE}
View(py$dat)
```

# Mutate

The *assign()* function appears to be a clean way to create new variables. 

```{python}
dat.assign(value = dat.points / dat.price).head(5)
```

# Summarize (Groupby)

Grouping and summarizing variables is straightforward as well. You can select which variables to group through the *groupby()* function and then perform different summary statistics. There is an *agg()* function to indicate which aggregations to perform. You can also use **numpy** functions directly like *sum()* to your grouped data. Lastly, you can use a dictionary style of input to summarize your data. I think this dictionary style format is very clean and come in handing when I need to perform various calculations on grouped data with different data types. 

```{python}
dat.groupby('country').agg('mean')
dat.groupby('country').sum()
dat.groupby('country').agg(['mean', 'sum']) # Multiple summary statistics
dat.groupby('country').agg({'price': 'mean'}) # With a dictionary
```

# Chaining 

Python does not appear to need actual operators to chain functions as R does with **%>%** from tidyverse's magrittr package. Python can take advantage of the dot notation and continually chain different functions as long as they are wrapped in their own parenthesis. 

```{python}
(dat.loc[:, ['country', 'variety', 'points']]  # Select columns
    .groupby(['country', 'variety'])           # Groupby
    .agg('mean')                               # Summary Statistic
    .query('country == "US"')                  # Filter for US Only
    .sort_values('points', ascending = False)  # Arrange in descending order
    .head(10))                                 # Show top 10  
```

