---
title: 'Learning Python: The Beginning'
author: ~
date: '2019-05-31'
slug: learning-python-the-beginning
categories: []
tags: ['Python', 'Select', 'Filter', 'TidyTuesday_Python']
description: ''
topics: []
---

```{r, setup, echo=FALSE}
# Set up options for rmarkdown chunks
knitr::opts_chunk$set(warning = FALSE, message = FALSE, comment = "#")
```

# Learning Python

I often see a debate online between Python and R as programming languages for *Data Science*. I am not interested in continuing that debate here, but rather just want to continue to learn other programming languages and have another tool available to use. I often read online Python is better for general use programming while R was developed with statisticians in mind. If that is the case I think it would be helpful to start learning Python and see where I can use it for my advantage.

In this post, I will just try to practice my typical R tidyverse workflow of:

1. Get Data
2. Load needed libraries
3. Select and filter columns. 

# Set up Python in Rmarkdown

The **reticulate** package helps integrate Python into your rmarkdown scripts. After loading the package, you are suggested to specify which environment to use Python. I am still confused on how exactly to do this but I found the following code to be most stable to start my script especially if I have to leave and come back later to finish it. 

```{r, setup-python}
library(reticulate)
use_virtualenv("r-reticulate")
```

## Load Modules and Data
```{python, load}
import pandas as pd
import numpy as np

url = 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-28/winemag-data-130k-v2.csv'
dat = pd.read_csv(url, index_col=0)
```

```{python, options}
# Stop Python from truncating number of columns
pd.set_option('display.max_columns', None)
pd.set_option('display.width', 1000)
pd.set_option('precision', 2)
```

# Previewing the Data

## View Data in Rstudio
```{r, view_data}
View(py$dat)
```

## Understanding the Big Picture of a Dataset

```{python, summary}
dat.info() # General Info (i.e. columns, non-null entries, data types)
dat.head(5) # View top 5 rows of the data
dat.price.describe() # Describe basic statistics for numeric columns
dat.country.describe() # Describe non-numeric (strings) columns 
```

The *describe()* function returns certain information depending on the column type selected (i.e. integers or strings). However, using *describe()* without selecting any column prior will run *describe()* on all numeric data types only.

# Selecting Columns 
```{python, select}
# How to find index of a column, using region_2 as example
dat.columns.get_loc('region_2')
dat.country.head(5) #Select column by name
dat['country'].head(5) #Alternative to select column by name
dat.iloc[:, 0]. head(5) #Select column by Index location
dat.iloc[:, [4, 6, 10]].head(5) # Select multiple columns
dat.iloc[:, 1:4].head(5)# Select slice of columns
dat.iloc[:, np.r_[2:4, 9:12]].head(5) # Combine Multiple and Slice
```

# Filter Rows
```{python, filter}
dat[dat['country'] == "US"].count() # Filtering String Columns
dat[dat['price'] > 500].count() # Filtering Strings
dat.query('price > 500').count() # Filter by query
dat.query('(price > 500) & (country == "US")').count() # Filter by more than 1 query
```